---
title: "ICD Sample File Generator"
author: "Seth Russell"
date: "10/5/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This script is used to create a sample data set of N rows with X columns of data. User can specify D columns of ICD diagnosis codes, P columns of ICD procedure codes, and G columns of extra data.

Either ICD9 or ICD10 codes can be generated.

### ICD 9
ICD 9 code list was obtained from CMS on October 2017 via: https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html - specific file downloaded was [Version 32 Full and Abbreviated Code Titles – Effective October 1, 2014 [ZIP, 1MB]](https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/Downloads/ICD-9-CM-v32-master-descriptions.zip) 

ICD 9 Files included in with this code are:

* CMS32_DESC_LONG_DX.txt - List of ICD9 diagnosis codes along with a long description
* CMS32_DESC_LONG_SG.txt - List of ICD9 procedure codes along with a long description

### ICD 10

ICD 10 code list was obtained from CMS on October 2017 via: https://www.cms.gov/Medicare/Coding/ICD10/2017-ICD-10-CM-and-GEMs.html - specific files downloaded were 
[2017 Code Descriptions in Tabular Order [ZIP, 2MB]](https://www.cms.gov/Medicare/Coding/ICD10/Downloads/2017-ICD10-Code-Descriptions.zip) and [2017 ICD-10-PCS Codes File [ZIP, 831KB]] (https://www.cms.gov/Medicare/Coding/ICD10/Downloads/2017-New-Procedure-Codes.zip)

ICD 10 Files included with this code are:

* icd10cm_codes_2017.txt - List of ICD10 diagnosis codes from "2017 2017 Code Descriptions in Tabular Order"
* icd10pcs_codes_2017.txt - List of ICD10 procedure codes from "2017 ICD-10-PCD Codes File"


```{r checkfiles}
library(tidyverse)
library(iotools)
library(data.table)
library(stringi)

files <- tibble(
  file_desc = c("icd9_dx", "icd9_pc", "icd10_dx", "icd10_pc"),
  file_name = c("data/CMS32_DESC_LONG_DX.txt",
                "data/CMS32_DESC_LONG_SG.txt",
                "data/icd10cm_codes_2017.txt",
                "data/icd10pcs_codes_2017.txt")
  )

col_names <- c("code", "description")

icd_data <- rbindlist(
  apply(files, 1, function(f) {
    if (file.exists(f["file_name"])) {
      # using microbenchmark tested data.table::fread, iotools::input.file, iotools:: read.csv.raw
      # found that iotools::input.file and data.table::fread were about the same (input.file just a tiny bit faster)
      # but since fread has better syntax, using fread.
      
      icd_version <- stri_extract_first_regex(f["file_desc"], "[0-9]+")
      code_type <- stri_extract_first_regex(f["file_desc"], "[a-z]+$")

      fread(f["file_name"],
            sep = "\t",
            header = FALSE,
            col.names = col_names,
            colClasses = c("character")) %>% add_column(icd_version = icd_version, code_type = code_type, .before = 1)
      }
    else {
      warning(paste(f["file_desc"], "file is missing - check current working directory is correct.\n"))
      # have to return emtpy data frame so as to no mess up rbindlist
      data.table(
        icd_version = character(),
        code_type =character(),
        code = character(),
        description = character()
      )
    }
  }
))

head(icd_data)
```
## Search for ICD codes

Now that the data is loaded into memory, do some validation or review of ICD codes. Here are a few examples:

```{r icdcodes}

search_all_cols = function(search_term) {
  apply(icd_data, 1, function(r) {
  any(stri_detect_regex(r, search_term))
  }) %>% subset(icd_data, .)
}

# search for string of characters anywhere
search_all_cols("516")

# search for L followed by 1 or more digits
search_all_cols("L\\d+")

search_for_code_pattern = function(v = NULL, search_term, ct = NULL) {
  dt <- icd_data
  
  if(!is.null(v))
    dt <- dt[icd_version == v]
  
  if(!is.null(ct))
    dt <- dt[code_type == ct]

  apply(dt, 1, function(r) {
    stri_detect_regex(r["code"], search_term)
  }) %>% subset(dt, .)
}

search_for_code_exact = function(v = NULL, search_term, ct = NULL) {
  dt <- icd_data
  
  if(!is.null(v))
    dt <- dt[icd_version == v]
  
  if(!is.null(ct))
    dt <- dt[code_type == ct]

  dt[code == search_term]
}

# example with all positional arguments
search_for_code_pattern(10, "^G318.*", "dx")

# just code to search for
search_for_code_pattern(search_term = 334)

# code and type (looking at both 9 & 10)
search_for_code_pattern(search_term = 334, ct = "pc")

# find exact matches
search_for_code_exact(search_term = "0010")


```

## Generate Random Data

Create a file using randomly selected ICD diagnosis and procedure codes

```{r genfile}
library(tictoc)

generate_sample <- function(v, n_rows, d_cols, p_cols, g_cols, pct_empty = 0.20) {

  tic("creating base data.tables")
  dx <- icd_data[icd_version == v & code_type == "dx"]
  pc <- icd_data[icd_version == v & code_type == "pc"]
  toc()
  
  tic("building indivual components")
  # this method for sample generation: 
  #   sample(dx$code, n_rows, replace=TRUE)
  # is about 35 times faster than
  #   dx[sample(1:nrow(dx), n_rows, replace=TRUE)]$code
  d <- as.data.table(replicate(d_cols, sample(dx$code, n_rows, replace=TRUE)))
  colnames(d) <- paste0('dx', 1:d_cols)
  p <- as.data.table(replicate(p_cols, sample(pc$code, n_rows, replace=TRUE)))
  colnames(p) <- paste0('pc', 1:p_cols)
  
  # hard coded range of 'fake' values to generate - probably should have this as a
  # separate function or input parameter
  g <- as.data.table(replicate(g_cols, sample(100000:999999, n_rows, replace=TRUE)))
  colnames(g) <- paste0('g', 1:g_cols)
  toc()
  
  tic("binding to one large data.table")
  # not sure why, but this line (which is how I saw on SO to combine multiple data.tables):
  #    as.data.table(do.call(cbind, c(d, p, g)))
  # runs in about 80 - 90 seconds for 1,000,000 rows
  # where as this line takes about 4 seconds for 1,000,000 rows
  #    as.data.table(cbind(cbind(d, p), g))
  # and this new method I discovered is even faster - about 0.6 second for 1,000,000 rows
  all_data <- copy(d)
  all_data[, colnames(p) := p]
  all_data[, colnames(g) := g]

  toc()
  
  tic("removing data at random")
  # using data.table set() method takes about 17 seconds for 1,000,000 rows
  for (j in 1:nrow(all_data)) {
    s <- sample(1:length(all_data), length(all_data) * pct_empty, replace = FALSE)
    set(all_data, i = j , s, NA) # integers using 'L' passed for efficiency
  }
  # # total time for sapply with 1,000,000 rows is about 75 seconds
  # # also updating row wise is about 2x faster than column wise update
  # if (pct_empty != 0) {
  #   all_data <- as.data.table(sapply(all_data, function(r) {
  #     r[(sample(1:length(r), length(r) * pct_empty, replace = FALSE))] <- NA
  #     r
  #   }))
  # }
  toc()
  
  tic("adding id column")
  # add an ID column - simple incremental number
  all_data[, id := seq_along(dx1), ]
  setcolorder(all_data, c("id", names(all_data)[1:(length(all_data)-1)]))
  toc()
  
  all_data
}

t <- generate_sample(
  v = 9,
  n_rows = 3407146, # KID file has 3407146 rows
  d_cols = 29,      # KID file has 29 columns of ICD diagnoses
  p_cols = 15,      # KID file has 15 columns of ICD procedures
  g_cols = 114      # KID file has 114 other columns
)

head(t)

library(feather)
# file is about 2.7GB for 3407146 rows of data
# takes about 6 seconds to write
# for this example, feather is slightly faster than fst for writing
tic("feather timing")
write_feather(t, "sample_icd9_data.feather")
toc()

# File is about 1.4GB for 3407146 rows of data
# takes about 210 seconds to write
#pccc_icd10_dataset <- t
#tic("rdata timing")
#save(pccc_icd10_dataset, file="pccc_icd10_dataset.rda")
#toc()

```

Example of running a benchmark to evaulate fastest option of function.

```{r benchmark}
library(microbenchmark)

mbm <- microbenchmark(
  floor(runif(100000000, min=0, max=100001)),
  sample(1:100000, 100000000, replace=TRUE),
  times=10
)
mbm


library(stringr)
library(stringi)
mbm <- microbenchmark(
  str_extract(files$file_desc, "[a-z]+$"),
  stri_extract_first_regex(files$file_desc, "[a-z]+$"),
  times=10000
)
mbm

```

